---
title: "Portfolio 1"
---

> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.

###In this portfolio, I will use the fish encounters data from tidyr
##First, I need to understand the data set, what's in it, what's measured, what do we have, and then think about what can be done

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
data(fish_encounters)

dim (fish_encounters)
cat("Number of observations:", nrow(fish_encounters), "\n")
cat("Number of variables:", ncol(fish_encounters), "\n")
#there are 114 observations of 3 variables (fish (fish identifier), station (measurement station), seen(was the fish seen? 1 if yes)) in the dataset

# Information about fish swimming down a river: each station represents an autonomous monitor that records if a tagged fish was seen at that location. Fish travel in one direction (migrating downstream). Information about misses is just as important as hits, but is not directly recorded in this form of the data.

```

#####I will first group the fish_encounters dataset by station to summarize the total number of encounters for each station. This allows us to compare the total fish encounters across different stations.
```{r fishencounters by station }
fish_encounters_summary <- fish_encounters %>%
  group_by(station) %>%
  summarize(
    count = n() # Count the number of entries for each station
  ) %>%
  arrange(desc(count))  # Arrange in descending order by total encounters

# View the summarized data
head(fish_encounters_summary)

# how many stations in total 
cat ("Number of stations:", nrow(fish_encounters_summary),"\n")

# the average number of fish encountered across stations 
average_fishencounter = nrow(fish_encounters)/nrow(fish_encounters_summary)
cat ("On average, there are", average_fishencounter, "fish encountered across stations.")

## barchart 
ggplot(data = fish_encounters, aes(x = station, y = seen)) +
  geom_bar(stat = "identity") +
  labs(x = "Station", y = "Number of Fish Encounters")

## there are 11 stations in the dataset. Release station and I80_1station enountered a total of 19 fish, which are the most. 
## on average, each station encountered 10 fish. 
```

#####I will group the dataset by fish id.
```{r fishencounters by id }
fish_encounters_summary2 <- fish_encounters %>%
  group_by(fish) %>%
  summarize(
    count = n() # Count the number of entries for each station
  ) %>%
  arrange(desc(count))  # Arrange in descending order by total encounters

# View the summarized data
head(fish_encounters_summary2)

# how many fish in total 
cat ("Number of fish:", nrow(fish_encounters_summary2),"\n")

#barchart
ggplot(data = fish_encounters, aes(x = fish, y = seen)) +
  geom_bar(stat = "identity") +
  labs(x = "Species", y = "Number of Fish Encounters")

## There are 19 fish in the data set. Across all stations, five fish were encountered the most, 11 times each.
```

